\textbf{ZKLoRA: Efficient Verification of LoRA-Base Model Compatibility for Large Language Models}

### 2025-01-13 paper/main.tex
update

\title{\textbf{zklora: Efficient Verification of LoRA-Base Model Compatibility for Large Language Models}}
\author{Anonymous Authors}
\date{\today}

@@

\begin{abstract}
Verifying that a private Low-Rank Adaptation (LoRA) module is truly compatible with a given large language model (LLM) can be daunting, especially when those LoRA parameters cannot be disclosed. 
We introduce \texttt{zklora}, a library for generating zero-knowledge proofs that demonstrate correct LoRA-base model compatibility \emph{without sharing LoRA weights}. 
A key insight is that the \emph{verification step} for each LoRA module remains on the order of \textbf{1--2 seconds}, enabling repeated or batch checking at scale. 
Our empirical evaluations confirm that \texttt{zklora} maintains this rapid per-module verification time even for multi-billion-parameter LLMs.
\end{abstract}

%------------------------------------------
% INTRO: emphasize verification challenge & clarify it's per-module
%------------------------------------------
\section{Introduction}
Large Language Models (LLMs) have attained remarkable success \cite{brown2020language, devlin2018bert}, but verifying fine-tuned modifications such as LoRA \cite{hu2021lora} can be difficult when the updated weights must remain private. 
Traditionally, one might re-run an entire forward pass or inspect thousands of parameters to ensure correctness, which is infeasible for massive models. 
\texttt{zklora} addresses this by generating a zero-knowledge proof of correctness for each LoRA module, guaranteeing that the private LoRA genuinely fits the public base model. 
Crucially, \textbf{the verification stage for each LoRA module} in \texttt{zklora} remains about 1--2 seconds, even at scales of multi-billion parameter base models. 
This rapid, per-module check allows the Base Model User to efficiently, independently verify the compatibility of the private LoRA parameters.

@@

We benchmarked \texttt{zklora} across several LLMs and smaller models with varying LoRA parameter counts. Our focus is verifying that, regardless of total parameter scale, \emph{the per-module verification step} remains swift. 
Figures~\ref{fig:verify}, \ref{fig:settings}, and \ref{fig:proof} illustrate how total LoRA parameters correlate with:
\begin{itemize}
    \item \textbf{Verification Time per Module} (Base Model User’s overhead),
    \item \textbf{Total Settings Time} (preparatory steps and key generation),
    \item \textbf{Total Proof Generation Time} (LoRA Owner’s overhead).
\end{itemize}

@@

Although proof generation may scale with parameter size, verification for each LoRA module is consistently on the order of seconds. We consider this \emph{fast, per-module verification} to be \texttt{zklora}’s core advantage in production LLM environments.

%------------------------------------------
% 2. Ensuring LoRA–Base Model Compatibility with zklora
%------------------------------------------
\section{Ensuring LoRA–Base Model Compatibility}

A typical workflow involves:
\begin{enumerate}
    \item \textbf{Base Model User} possessing a large LLM. 
    \item \textbf{LoRA Owner} having proprietary low-rank adapter weights. 
\end{enumerate}
However, these two parties must confirm the LoRA truly “fits” the base model. Using \texttt{zklora}, the LoRA Owner proves the adapter’s compatibility \emph{without} exposing private weights:

@@

\node[block, label=above:\textbf{Input}] (input) {$\mathbf{x}$};
\node[block, right=2.0cm of input] (W1) {$\mathbf{W}_1$};
\node[dashedblock, above=1.6cm of W1, label=above:\textbf{LoRA Owner}] (LoRA) {LoRA for Layer 1};
\node[block, right=2.0cm of W1] (act) {ReLU};
\node[block, right=2.0cm of act] (W2) {$\mathbf{W}_2$};
\node[block, label=above:\textbf{Output}, right=2.0cm of W2] (output) {$\mathbf{y}$};

\draw[->] (input) -- node[above, font=\scriptsize]{local multiply} (W1);
\draw[->, dashed] (W1.north) to[bend left=15] node[left, font=\scriptsize]{partial acts} (LoRA.south);
\draw[->, dashed] (LoRA.south) to[bend left=15] node[right, font=\scriptsize]{updated $\mathbf{z}'$} (W1.north);
\draw[->] (W1) -- node[above, font=\scriptsize]{$\mathbf{z}'$} (act);
\draw[->] (act) -- node[above, font=\scriptsize]{hidden} (W2);
\draw[->] (W2) -- (output);

\node[draw, rounded corners, thick, color=blue!40!black,
      fit=(input)(W1)(act)(W2)(output),
      label={[font=\small, color=blue!40!black]above:\textbf{Base Model User}},
      inner sep=0.4cm
] (baseModelFit) {};
\end{tikzpicture}
\caption{zklora inference flow on a 2-layer neural network where the first layer has private LoRA weights added.}
\label{fig:toy-mlp-lora}
\end{figure}

@@

\texttt{zklora} compiles an ONNX model (with LoRA included) into a constraint circuit. The user’s partial activations feed the circuit as inputs:

@@

\paragraph{(5) Verification per Module.}
Each LoRA module’s proof is tested with the verification key, taking \textbf{1--2 seconds}. This minimal overhead is \texttt{zklora}’s main advantage for frequent or large-scale checks.

@@

\texttt{zklora} provides a fast, robust mechanism to ensure that private LoRA modules remain consistent with a large base model. Our evaluations indicate that each LoRA module’s correctness can be verified in about 1--2 seconds, even for multi-billion-parameter LLMs. This efficiency bridges the gap between privacy-preserving LoRA development and practical, real-time validation in large-scale deployments. Future expansions could integrate multi-owner LoRAs, advanced zero-knowledge proofs for further performance gains, or partial data-privacy frameworks to shield user inputs as well as LoRA parameters.

>>>>

\title{\textbf{ZKLoRA: Efficient Verification of LoRA-Base Model Compatibility for Large Language Models}}
\author{
    \textbf{Bidhan Roy, Peter Potash, Marcos Villagra} \\
    Bagel Research Team\footnotemark[1] \\
    \texttt{bidhan@bagel.net, peter@bagel.net, marcos@bagel.net}
}
\date{\today}

@@

\footnotetext[1]{Bagel is a research lab, making open source AI monetizable.}

\begin{abstract}
Verifying that a private Low-Rank Adaptation (LoRA) module is truly compatible with a given large language model (LLM) can be daunting, especially when those LoRA parameters cannot be disclosed. 
We introduce \texttt{ZKLoRA}, a library for generating zero-knowledge proofs that demonstrate correct LoRA-base model compatibility \emph{without sharing LoRA weights}. 
A key insight is that the \emph{verification step} for each LoRA module remains on the order of \textbf{1--2 seconds}, enabling repeated or batch checking at scale. 
Our empirical evaluations confirm that \texttt{ZKLoRA} maintains this rapid per-module verification time even for multi-billion-parameter LLMs.
\end{abstract}

%------------------------------------------
% INTRO: emphasize verification challenge & clarify it's per-module
%------------------------------------------
\section{Introduction}
Large Language Models (LLMs) have attained remarkable success \cite{brown2020language, devlin2018bert}, but verifying fine-tuned modifications such as LoRA \cite{hu2021lora} can be difficult when the updated weights must remain private. 
Traditionally, one might re-run an entire forward pass or inspect thousands of parameters to ensure correctness, which is infeasible for massive models. 
\texttt{ZKLoRA} addresses this by generating a zero-knowledge proof of correctness for each LoRA module, guaranteeing that the private LoRA genuinely fits the public base model. 
Crucially, \textbf{the verification stage for each LoRA module} in \texttt{ZKLoRA} remains about 1--2 seconds, even at scales of multi-billion parameter base models. 
This rapid, per-module check allows the Base Model User to efficiently, independently verify the compatibility of the private LoRA parameters.

@@

We benchmarked \texttt{ZKLoRA} across several LLMs and smaller models with varying LoRA parameter counts. Our focus is verifying that, regardless of total parameter scale, \emph{the per-module verification step} remains swift. 
Figures~\ref{fig:verify}, \ref{fig:settings}, and \ref{fig:proof} illustrate how total LoRA parameters correlate with:
\begin{itemize}
    \item \textbf{Verification Time per Module} (Base Model User’s overhead),
    \item \textbf{Total Settings Time} (preparatory steps and key generation),
    \item \textbf{Total Proof Generation Time} (LoRA Owner’s overhead).
\end{itemize}

@@

Although proof generation may scale with parameter size, verification for each LoRA module is consistently on the order of seconds. We consider this \emph{fast, per-module verification} to be \texttt{ZKLoRA}’s core advantage in production LLM environments.

%------------------------------------------
% 2. Ensuring LoRA–Base Model Compatibility with ZKLoRA
%------------------------------------------
\section{Ensuring LoRA–Base Model Compatibility}

A typical workflow involves:
\begin{enumerate}
    \item \textbf{Base Model User} possessing a large LLM. 
    \item \textbf{LoRA Owner} having proprietary low-rank adapter weights. 
\end{enumerate}
However, these two parties must confirm the LoRA truly “fits” the base model. Using \texttt{ZKLoRA}, the LoRA Owner proves the adapter’s compatibility \emph{without} exposing private weights:

@@

% Base Model User components
\node[block, label=above:\textbf{Input}] (input) {$\mathbf{x}$};
\node[block, right=2.0cm of input] (W1) {$\mathbf{W}_1$};
\node[block, right=2.0cm of W1] (act) {ReLU};
\node[block, right=2.0cm of act] (W2) {$\mathbf{W}_2$};
\node[block, label=above:\textbf{Output}, right=2.0cm of W2] (output) {$\mathbf{y}$};

% LoRA Owner components
\node[dashedblock, above=1.6cm of W1, label=above:\textbf{LoRA Owner}] (LoRA) {LoRA for Layer 1};
\node[dashedblock, above=1.6cm of act] (proofGen) {Proof Generation};
\node[dashedblock, above=1.6cm of W2] (zkProof) {Zero-Knowledge Proof};

% Arrows for Base Model User
\draw[->] (input) -- node[above, font=\scriptsize]{local multiply} (W1);
\draw[->] (W1) -- node[above, font=\scriptsize]{$\mathbf{z}'$} (act);
\draw[->] (act) -- node[above, font=\scriptsize]{hidden} (W2);
\draw[->] (W2) -- (output);

% Arrows for LoRA Owner
\draw[->, dashed] (W1.north) to[bend left=15] node[left, font=\scriptsize]{partial acts} (LoRA.south);
\draw[->, dashed] (LoRA.south) to[bend left=15] node[right, font=\scriptsize]{updated $\mathbf{z}'$} (W1.north);
\draw[->, dashed] (LoRA) -- (proofGen);
\draw[->, dashed] (proofGen) -- (zkProof);

% Verification process
\node[below=0.5cm of W2, font=\footnotesize, align=center] (verify) {Verification\\(1--2 sec)};
\draw[->, dashed] (zkProof) -- (verify);

% Base Model User boundary
\node[draw, rounded corners, thick, color=blue!40!black,
      fit=(input)(W1)(act)(W2)(output),
      label={[font=\small, color=blue!40!black]above:\textbf{Base Model User}},
      inner sep=0.4cm
] (baseModelFit) {};

% Additional annotations
\node[above=0.5cm of zkProof, font=\footnotesize, align=center] (zkProofGen) {Zero-Knowledge\\Proof Generation};
\draw[->, dashed] (zkProof) -- (zkProofGen);

\end{tikzpicture}
\caption{ZKLoRA inference flow on a 2-layer neural network where the first layer has private LoRA weights added. The LoRA Owner generates a zero-knowledge proof to ensure compatibility, and the Base Model User verifies this proof quickly.}
\label{fig:toy-mlp-lora}
\end{figure}

@@

\texttt{ZKLoRA} compiles an ONNX model (with LoRA included) into a constraint circuit. The user’s partial activations feed the circuit as inputs:

@@

\paragraph{(5) Verification per Module.}
Each LoRA module’s proof is tested with the verification key, taking \textbf{1--2 seconds}. This minimal overhead is \texttt{ZKLoRA}’s main advantage for frequent or large-scale checks.

@@

\texttt{ZKLoRA} provides a fast, robust mechanism to ensure that private LoRA modules remain consistent with a large base model. Our evaluations indicate that each LoRA module’s correctness can be verified in about 1--2 seconds, even for multi-billion-parameter LLMs. This efficiency bridges the gap between privacy-preserving LoRA development and practical, real-time validation in large-scale deployments. Future expansions could integrate multi-owner LoRAs, advanced zero-knowledge proofs for further performance gains, or partial data-privacy frameworks to shield user inputs as well as LoRA parameters.
